---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data <- read.csv(unzip("activity.zip"))
```

It is a simple data frame with date, steps and interval variables. 

```{r}
str(data)
```

## What is mean total number of steps taken per day?

Now I will group the dataframe by date

```{r}
g <- group_by(data, date)
```

Here I save the intermediate result into the object *g*. Now I'm ready to extract useful values.

```{r}
daily_steps <- summarise(g, sum=sum(steps))
```

Let's look to the total number of steps taken per day
```{r}
hist(daily_steps$sum)
```
Most of the days the number of step was between 10000 and 15000.

I have a kind of problem with NA values.
```{r}
summary(daily_steps$sum)
```


I'll just drop the days with NA values.

```{r}
good_daily_steps <- daily_steps[!is.na(daily_steps$sum),]
```

Finally, we're ready to get some statistics about steps taken per day, including mean and median values. 

```{r}
summary(good_daily_steps$sum)
```


Mean is a little bit more than 10000 steps. 

Median is about the same value (just less by 10)

Done.


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
